---
title: "Line Street Maps - OSM Data"
author: "Will Ball"
date: "22/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


``` {r setup, echo = FALSE}
library(osmdata)
library(tidyverse)
library(beepr)
```

## Line Street Maps

This document explains how to create line maps using OSM data and ggplot2. First, we collect Open Street Map data on main streets, small streets and any rivers that we want to plot. This uses the osmdata package.

```{r Data}
area <- "Aberdeen, Scotland" # Insert city/area name here within speech marks

streets <- getbb(area) %>%
  opq() %>%
  add_osm_feature(key = "highway", 
                  value = c("motorway", "primary", 
                            "secondary", "tertiary")) %>%
  osmdata_sf()

small_streets <- getbb(area) %>%
  opq() %>%
  add_osm_feature(key = "highway", 
                  value = c("residential", "living_street",
                            "unclassified",
                            "service", "footway")) %>%
  osmdata_sf()

river <- getbb(area) %>%
  opq() %>%
  add_osm_feature(key = "waterway", value = "river") %>%
  osmdata_sf()

a <- getbb(area) # Creates an object that lists the boundary box limits for the area
beep("coin")
```

## Plotting

Here we plot the data extracted above. Each one forms a 'geom_sf'. We can specify the colour, size and opacity (alpha).

```{r plot}
(p <- ggplot() +
  geom_sf(data = streets$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .3,
          alpha = .6) +
  geom_sf(data = small_streets$osm_lines,
          inherit.aes = FALSE,
          color = "black",
          size = .2,
          alpha = .4) +
  geom_sf(data = river$osm_lines,
          inherit.aes = FALSE,
          color = "light blue",
          size = .3,
          alpha = .6) +
  coord_sf(xlim = c(a[1], a[3]), 
           ylim = c(a[2], a[4]),
           expand = FALSE) +
    theme_void()
)
beep("coin")
```

## Save

Now to save the plot. Don't forget to change the filename

``` {r save}
ggsave(filename = "Aberdeen.png", plot = p, height = 3, width = 3, units = "in", dpi = 500)

beep("coin")
```
